---
title: "p8105 hw2"
author: "Yue Chen"
date: "9/24/2020"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
library(dplyr)
```

## Problem 1

Read the Mr. Trashwheel dataset.
```{r}
trash_wheel = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx") %>%
   janitor::clean_names() %>%
   select(-x15, -x16, -x17) %>%
   drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2017.
```{r}
prcip_2017 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 6, skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017)
```

Read precipitation data for 2018.
```{r}
prcip_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5, skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018)
```

Combine precipitation datasets and convert month to a character variable.
```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

prcip_1718 = 
  bind_rows(prcip_2017, prcip_2018)

left_join(prcip_1718, month_df, by = "month")
```

The dataset contains information of a water-wheel vessel that removes trash from the Inner Harbor in Baltimore, recording month, year, type of transh each dumpster collects, and weight and volume of trash collected. There are `r nrow(trash_wheel)` obeservations in final dataset.
The total precipitation in 2018 was `r sum(prcip_2018, pull(prcip_2018, total))`. The median number of sports balls in a dumpster in 2017 was `r median(filter(trash_wheel, pull(trash_wheel, year) == 2017)$sports_balls)`.

## Problem 2

```{r}
nyc_transit = read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  dplyr::select(line:route11, entry, vending, entrance_type, ada)
  mutate(nyc_transit,
   entry = recode(entry, 'YES' = TRUE, 'NO' = FALSE)
  )
```

The dataset contains line, station name, station latitude/longitude, route, entry, entrance type, vending, ADA
compliance of each subway station in NYC. So far, I have cleaned variable names, removed several variables, 
and converted entry variable from character to logical variable. There are `r ncol(nyc_transit)` variables and `r nrow(nyc_transit)` observations. These data are not tidy.

There are `r count(distinct(nyc_transit, station_name, line))` distinct stations.

There are `r sum(distinct(nyc_transit, station_name, line, ada)$ada, na.rm = TRUE)` stations that are ADA compliant.

The proportion of station entrances without vending allow entrance is `r nrow(filter(nyc_transit, vending == "NO", entry == 'YES'))/nrow(filter(nyc_transit, vending == "NO"))`.

Reformat data:
```{r}
nyc_transit_tidy = 
  gather(
    nyc_transit, key = "route_number", value = "route_name", route1:route11
  )
```



There are `r count(distinct(filter(nyc_transit_tidy, route_name == 'A')))` distinct stations serve the A train.

Of the stations that serve the A train, there are `r count(distinct(filter(nyc_transit_tidy, route_name == 'A', ada == TRUE)))` stations that are ADA compliant.

## Problem 3


