---
title: "p8105 hw2"
author: "Yue Chen"
date: "9/24/2020"
output: github_document
---

```{r}
library(tidyverse)
library(readxl)
```

Problem 1

Read the Mr. Trashwheel dataset.
```{r}
trash_wheel = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx") %>%
   janitor::clean_names() %>%
   select(-x15, -x16, -x17) %>%
   drop_na(dumpster) %>%
  mutate(
    sports_balls = round(sports_balls),
    sports_balls = as.integer(sports_balls)
  )
```

Read precipitation data for 2017.
```{r}
prcip_2017 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 6, skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2017)
```

Read precipitation data for 2018.
```{r}
prcip_2018 = read_excel("./data/Trash-Wheel-Collection-Totals-8-6-19.xlsx", sheet = 5, skip = 1) %>%
  janitor::clean_names() %>%
  drop_na(month) %>%
  mutate(year = 2018)
```

Combine precipitation datasets and convert month to a character variable.
```{r}
month_df = 
  tibble(
    month = 1:12,
    month_name = month.name
  )

prcip_1718 = 
  bind_rows(prcip_2017, prcip_2018)

left_join(prcip_1718, month_df, by = "month")
```

The dataset contains information of a water-wheel vessel that removes trash from the Inner Harbor in Baltimore, recording month, year, type of transh each dumpster collects, and weight and volume of trash collected. There are `r nrow(trash_wheel)` obeservations in final dataset.
The total precipitation in 2018 was `r sum(prcip_2018, pull(prcip_2018, total))`. The median number of sports balls in a dumpster in 2017 was `r median(filter(trash_wheel, pull(trash_wheel, year) == 2017)$sports_balls)`.






